Black swans in ecological time series\\Black swans in population dynamics\\Heavy-tailed ecological time series\\Black swans in ecological dynamics\\Black swan ecological dynamics\\ \ldots

### 2014-10-17 supp.tex
Trivial edits

\begin{enumerate}
\item
  To remove populations with unreliable population indices that could be strongly confounded with economics and sampling effort, we removed all populations with a sampling protocol listed as \texttt{harvest} as well populations with the words \texttt{harvest} or \texttt{fur} in the cited reference title.
\item
  We removed all populations with uneven sampling intervals. I.e.~we removed populations that didn't have a constant difference between the ``decimal year begin'' and ``decimal year end'' columns.
\item
  We removed all populations rated as $< 2$ in the GPDD quality assessment (on a scale of $1$ to $5$, with $1$ being the lowest quality data) \citep[following][]{sibly2005, ziebarth2010}
\item
  Populations with negative abundance values were assumed to be log values and transformed by taking $10$ to the power of the recorded abundance. In many cases this was noted for the population, but not in all cases. We inspected each of these \totalAssumedLog~time series to make sure our assumption made sense.
\item
  We filled in all missing time steps with \texttt{NA} values and imputed single missing values with the geometric mean of the previous and following values. We chose a geometric mean to be linear on the log scale that the Gompertz and Ricker-logistic models were fitted on.
\item
  We filled in single recorded values of zero with the lowest non-zero value in the time series \citep[following][]{brook2006a}. This assumes that single values of zero result from abundance being low enough that censusing missed present individuals. We turned multiple zero values in a row into \texttt{NA} values. This implies that multiple zero values were either censusing errors or caused by emigration. Regardless, our population models were fitted on a multiplicative (log) scale and so could not account for zero abundance.
\item
  We removed all populations with four or more identical values in a row since these suggest either recording error or extrapolation between two observations.
\item
  We removed all populations without at least four unique values \citep[following][]{brook2006a}.
\item
  We then wrote an algorithm to find the longest unbroken window of abundance (no \texttt{NA}s) with at least 20 time steps in each population time series. If there were any populations with multiple windows of identical length, we took the most recent window. This is a longer window than used in some previous analyses \citep[e.g.][]{brook2006a}, but since our model attempts to capture the shape of the distribution tails, our model requires more data.
\item
  Finally, we removed GPDD Main ID \texttt{20531}, which we noticed was a duplicate of \texttt{10139} (a heron population).
\end{enumerate}
\noindent
We provide a supplemental figure of all the time series included in our analysis and indicate which values were interpolated (\percImputedPops\% of populations had at least one point interpolated and only \percImputedPoints\% of the total observations were interpolated) (Fig.~\ref{fig:all-ts}). Table S1 shows the final taxonomic breakdown and the number of populations with interpolated values.

>>>>

\begin{enumerate}
\item
  To remove populations with unreliable population indices that could be strongly confounded with economics and sampling effort, we removed all populations with a sampling protocol listed as \texttt{harvest} as well populations with the words \texttt{harvest} or \texttt{fur} in the cited reference title.
\item
  We removed all populations with uneven sampling intervals. I.e.~we removed populations that didn't have a constant difference between the ``decimal year begin'' and ``decimal year end'' columns.
\item
  We removed all populations rated as $< 2$ in the GPDD quality assessment (on a scale of $1$ to $5$, with $1$ being the lowest quality data) \citep[following][]{sibly2005, ziebarth2010}
\item
  Populations with negative abundance values were assumed to be log values and transformed by taking $10$ to the power of the recorded abundance. In many cases this was noted for the population, but not in all cases. We inspected each of these \totalAssumedLog~time series to make sure our assumption made sense.
\item
  We filled in all missing time steps with \texttt{NA} values and imputed single missing values with the geometric mean of the previous and following values. We chose a geometric mean to be linear on the log scale that the Gompertz and Ricker-logistic models were fitted on.
\item
  We filled in single recorded values of zero with the lowest non-zero value in the time series \citep[following][]{brook2006a}. This assumes that single values of zero result from abundance being low enough that censusing missed present individuals. We turned multiple zero values in a row into \texttt{NA} values. This implies that multiple zero values were either censusing errors or caused by emigration. Regardless, our population models were fitted on a multiplicative (log) scale and so could not account for zero abundance.
\item
  We removed all populations with four or more identical values in a row since these suggest either recording error or extrapolation between two observations.
\item
  We removed all populations without at least four unique values \citep[following][]{brook2006a}.
\item
  We then wrote an algorithm to find the longest unbroken window of abundance (no \texttt{NA}s) with at least $20$ time steps in each population time series. If there were any populations with multiple windows of identical length, we took the most recent window. This is a longer window than used in some previous analyses \citep[e.g.][]{brook2006a}, but since our model attempts to capture the shape of the distribution tails, our model requires more data.
\item
  Finally, we removed GPDD Main ID \texttt{20531}, which we noticed was a duplicate of \texttt{10139} (a heron population).
\end{enumerate}
\noindent
We provide a supplemental figure of all the time series included in our analysis and indicate which values were interpolated (\percImputedPops\% of populations had at least one point interpolated and only \percImputedPoints\% of the total observations were interpolated) (Fig.~\ref{fig:all-ts}). Table S1 shows the final taxonomic breakdown and the number of populations with interpolated values.
